### 2020-04-01 05-translation.tex
Update 05-translation.tex

\begin{definition}[Translation]
\label{def:trans}
Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
$$
\llbracket{\cal A},\gamma \rrbracket 
\triangleq
 \restrict{\mathit{names}}(I~|~\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1)
$$
where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}\, \cup\, \{cxt\}$, with  $\mathit{reacts}$ be the set of reaction names $r_j $, $\mathit{cxts}$ the set of context names $\{\mathit{cxt}\}$,  $\mathit{ents}$ the set of decorated entity names  $\{s_i,s_o,\widehat{s}_i,\widehat{s}_o,\widetilde{s}_i,\widetilde{s}_o,\overline{s}_i,\overline{s}_o,\underline{s}_i,\underline{s}_o| s \in S\}$, and  $\mathit{prods}$ be the set of names for production $p_j$ associated to each reaction. 
%For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
{\color{red} For technical reasons, we introduce the init process $I \defeq \link{\silent}{r_1}.I$
to allow the name $r_1$ to be matched; also, 
for notational convenience, we fix that 
%$r_1 = \silent$,
$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u$ the number of reacts.}
\end{definition}

@@

 \begin{restatable}[]{lemma}{lemmastruct}
 \label{lem:struct}
%\begin{lemma}
 %\label{lem:struct}
  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation. If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{\mathit{names}}\upsilon} P')$ is a transition of $P$, then
  \begin{enumerate}
  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s \in S$, the corresponding channel $s$ (suitably decorated) appear in $\upsilon$; 
  the channel $\mathit{cxt}$ appears in $\upsilon$;
  %processes $P_s$ (or $\overline{P_s}$) 
%  for each  $Cxt_s$ are
 %involved in the transition $t$;
  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}^n$ is  overlapped by exactly one solid link offered by  processes representing entities.
  %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.

@@

{\color{red} \begin{example}\label{ex:backbone}
 Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and the reactions $r_1=(s1,\ , s2)$ and $r_2 = (s2, \ , s1)$ that produce $s2$ if $s1$ is present and $s1$ if $s2$ is present. For simplicity we do not consider inhibitors. Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$.
 %, for the first $n$ steps.
 Our translation follows:\\% includes the processes:\\
 \[
 \begin{array}{l@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}ll@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}l}
 P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\tilde{s2}_i}{\noact}\chainedlink{\noact}{\tilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots; & 
   P_{s2} &\defeq& \startchain{\hat{s2}_i}\chainedlink{\hat{s2}_o}{\noact}
   \chainedlink{\noact} {\tilde{s2}_i}\chainend{\tilde{s2}_o}.P_{s2} + \dots;\\
  P_{r_2} &\defeq& \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{cxt}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; &
 P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}
 %{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}
 .P_{s1} + \dots; \\
 Cxt& \defeq & \startchain{cxt}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}
  \chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}.Cxt\\

  %Cxt& \defeq & \startchain{cxt}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}
  %\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}.Cxt
 %Cxt_{s1}& \defeq & \startchain{cxt_1}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}\chainend{cxt_2}.Cxt_{s1} &
 % Cxt _{s2} &\defeq& \startchain{cxt_2}\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}. Cxt_{s2}
\end{array}
\]  \\
   Now, we assume that $s1$ only is in the initial state of ${\cal A}$; for clarity we partially specify the {\tt link}-calculus code of the processes. In Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of a transition of the c\CNA \ system: $\restrict{names} (I | P_{s1}|P_{s2}| P_a|\dots|Cxt)$. The yellow blocks are referred to init process $I$, and the processes encoding the reactions, $P_{r_1}$ and $P_{r_2}$, and the context $Cxt$. As the figure puts in evidence, these two kinds of processes determine the structure of the link chain, from end to end, i.e. from the left $\silent$ to the right one. We could say that these processes form the \emph{backbone} of the interaction.
   In contrast, the processes encoding the entities, $P_{s1}$,  $P_{s2}$, provides the
    solid links to overlap the virtual links   of the backbone.

@@

 \draw (10.22,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\silent $};
%{${\scriptscriptstyle \tilde{s1}_i}$};
 \draw (9.74,0.745) node [minimum size=0.485cm,draw,fill=yellow,draw=yellow]{$\ $};
 \draw (9.735,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
 %  \draw (10.8,0.5) node{$\dots$};
% \draw (10,1) -- (10,0.5) -- (10.5,0.5)--(10.5,0);
  %\draw (10.31,0.75)  node{$\noact$};
%  \draw (10.75,0.75)  node {$\ $};
%  \draw (10.75,0.26)  node{$\ $};
% \draw (11.25,0.25)  node {$\noact$};
%   \draw (11,1) -- (11,0.5) -- (11.5,0.5) -- (11.5,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
%    \draw (11,1) -- (11.5,1) -- (12,1) -- (12,0.5) -- (12.5,0.5)--(12.5,0)--(11.5,0);
   %\draw (11.25,0.25)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {$\noact$};
%    \draw (11.31,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_o}$};
%    \draw (11.75,0.26)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
%      \draw (11.75,0.75)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
%     \draw (12.2,0.25) node[minimum size=0.478cm,draw,fill=yellow,draw=yellow]{\;$\silent$\;  };
  \end{tikzpicture}
  \caption{The link chain structure arising from reactions and context processes.}
   \label{fig:backbone}
 \end{figure}
 \end{example}
}

>>>>

In the following use the following conventions for denoting different categories of names:
\begin{itemize}
\item
$\mathit{decs} \defeq \{ s, \overline{s}, \widetilde{s},\widehat{s},\underline{s}~|~ s\in S\}$ is the set of names for decorated entities;
\item
$\mathit{ents} \defeq \{ d_i, d_o~|~ d\in \mathit{decs}\}$ is the set names for entities;
\item
$\mathit{reacts}\defeq \{r_1,...,r_u\}$ is the set of reaction names associated with each reaction $aj$;
\item 
$\mathit{prods}\defeq \{p_1,...,p_u\}$ is the set of names for production $p_j$, associated with  reaction $aj$.
\end{itemize}

\begin{definition}[Translation]
\label{def:trans}
Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
$$
\llbracket{\cal A},\gamma \rrbracket 
\triangleq
 \restrict{\mathit{names}}(I~|~\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1)
$$
where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}\, \cup\, \{cxt\}$. 
%For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
For technical reasons, we introduce the init process $I \defeq \link{\silent}{r_1}.I$
to allow the name $r_1$ to be matched at the start of any chain; also, 
for notational convenience, we fix that 
%$r_1 = \silent$,
$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u$ the cardinality of $\mathit{reacts}$.
\end{definition}

@@

 \begin{restatable}[]{lemma}{lemmastruct}
 \label{lem:struct}
%\begin{lemma}
 %\label{lem:struct}
  Let ${\cal A} = (S, A)$ be a RS and let $\pi=(\gamma,\delta)$ be an extended  interactive process in ${\cal A}$. Let $P=\llbracket{\cal A},\gamma\rrbracket$ its c\CNA~translation. If $exists$ $P'$ such that $t= (P \xrightarrow{\restrict{\mathit{names}}\upsilon} P')$ is a transition of $P$, then
  \begin{enumerate}
  \item for each reaction  $a_j \in A$, the corresponding channels $r_j$ and $p_j$ appear in  $\upsilon$; for each entity $s \in S$, the corresponding channel $s$ (suitably decorated) appear in $\upsilon$; 
  the channel $\mathit{cxt}$ appears in $\upsilon$;
  %processes $P_s$ (or $\overline{P_s}$) 
%  for each  $Cxt_s$ are
 %involved in the transition $t$;
  \item for each reaction $a \in A$ and each entity $s \in S$, each virtual link offered by processes $P_a$ and $\mathit{Cxt}^1$ is  overlapped by exactly one solid link offered by  processes representing entities.
  %each process $P_s $ (or its alternative version $\overline{P}_s$), with $s \in S$, participates at most  with one solid link for each process $P_a$, and with exactly one solid link for each process $Cxt_s$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.

@@

\begin{example}\label{ex:backbone}
 Let ${\cal A}$ be a RS whose specification contains two entities, $s1$ and $s2$, and the reactions $r_1=(s1,\ , s2)$ and $r_2 = (s2, \ , s1)$ that produce $s2$ if $s1$ is present and $s1$ if $s2$ is present. For simplicity we do not consider inhibitors, but the reader can assume a void inhibitor is present in both reactions. Then, we assume an  extended interactive process $\pi=(\gamma,\delta)$ where the context $\gamma$ provides  $s1$ and $s2$ at every step, but we assume that only $s1$ is initially present. The corresponding c\CNA \ process is: 
$$\restrict{\mathit{names}} (~I ~|~ P_{s1}~|~\overline{P}_{s2}~|~ P_{r1}~|~P_{r2}~|~\mathit{Cxt})$$
 where:
 \[
 \begin{array}{l@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}ll@{\hspace{-0.009cm}}c@{\hspace{-0.009cm}}l}
 P_{r_1} &\defeq& \startchain{r_1}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{r_2}{\noact} \chainedlink{\noact}{p_1}\chainedlink{\widetilde{s2}_i}{\noact}\chainedlink{\noact}{\widetilde{s2}_o}\chainend{p_2}.P_{r_1} + \dots; & 
   P_{s1} &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\widehat{s1}_i}\chainend{\widehat{s1}_o}
 %{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}
 .P_{s1} + \dots; \\   
  P_{r_2} &\defeq& \dots + \startchain{r_2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{cxt}{\noact} \chainedlink{\noact}{p_2}\chainend{\silent}.P_{r_2} + \dots; &
  \overline{P}_{s2} &\defeq& \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}
   \chainedlink{\noact} {\widetilde{s2}_i}\chainend{\widetilde{s2}_o}.P_{s2} + \dots;\\
 Cxt& \defeq & \startchain{cxt}\chainedlink{\widehat{s1}_i}{\noact}\chainedlink{\noact}{\widehat{s1}_o}
  \chainedlink{\widehat{s2}_i}{\noact}\chainedlink{\noact}{\widehat{s2}_o}\chainend{p_1}.Cxt\\

  %Cxt& \defeq & \startchain{cxt}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}
  %\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}.Cxt
 %Cxt_{s1}& \defeq & \startchain{cxt_1}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1}_o}\chainend{cxt_2}.Cxt_{s1} &
 % Cxt _{s2} &\defeq& \startchain{cxt_2}\chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainend{p_1}. Cxt_{s2}
\end{array}
\]  \\
For clarity we show the code of the processes just in part, to focus on the prefixes that will be involved in the first transition of the system. In Figure~\ref{fig:backbone} we show the  structure of a link chain label related to the execution of such a transition. The yellow blocks are referred to init process $I$, to the processes encoding the reactions, $P_{r_1}$ and $P_{r_2}$, and to the context $\mathit{Cxt}$. As the figure puts in evidence, these two kinds of processes determine the structure of the link chain, from end to end, i.e. from the left $\silent$ to the right one. We could say that these processes form the \emph{backbone} of the interaction.
   In contrast, the processes encoding the entities, $P_{s1}$,  $\overline{P}_{s2}$, provides the
    solid links to be merged with the virtual links of the backbone (i.e. to be plugged in the backbone).

@@

 \draw (10.22,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\silent $};
%{${\scriptscriptstyle \tilde{s1}_i}$};
 \draw (9.74,0.745) node [minimum size=0.485cm,draw,fill=yellow,draw=yellow]{$\ $};
 \draw (9.735,0.26) node [minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
 %  \draw (10.8,0.5) node{$\dots$};
% \draw (10,1) -- (10,0.5) -- (10.5,0.5)--(10.5,0);
  %\draw (10.31,0.75)  node{$\noact$};
%  \draw (10.75,0.75)  node {$\ $};
%  \draw (10.75,0.26)  node{$\ $};
% \draw (11.25,0.25)  node {$\noact$};
%   \draw (11,1) -- (11,0.5) -- (11.5,0.5) -- (11.5,0);% -- (10,0) -- (9,0) -- (9,0.5) -- (8.5,0.5) -- (8.5,1);
%    \draw (11,1) -- (11.5,1) -- (12,1) -- (12,0.5) -- (12.5,0.5)--(12.5,0)--(11.5,0);
   %\draw (11.25,0.25)  node[minimum size=0.48cm,draw,fill=yellow,draw=yellow] {$\noact$};
%    \draw (11.31,0.75) node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{${\scriptscriptstyle \tilde{s1}_o}$};
%    \draw (11.75,0.26)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
%      \draw (11.75,0.75)node[minimum size=0.48cm,draw,fill=yellow,draw=yellow]{$\ $};
%     \draw (12.2,0.25) node[minimum size=0.478cm,draw,fill=yellow,draw=yellow]{\;$\silent$\;  };
  \end{tikzpicture}
  \caption{The link chain structure arising from reactions and context processes.}
   \label{fig:backbone}
 \end{figure}
 \end{example}
